A continuous urbanization index to assess phylogenetic adaptations in urban bird living

### 2018-05-15 ms.Rmd
Some polishing/minor changes of results

## Taxonomic considerations
Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed. The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/]. There were minor differences which existed in common names, scientific names, and recognized species among the three datasets. A taxonomic look-up table was made which combined the three different taxonomies (Appendix 2). Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset. If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher). Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.

# RESULTS
A total of 637,482 checklists provided a total of 5,944,819 observations of 580 different species which had > 100 observations. The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615). The mean number of observations for a species was 10,250. Density plots of all 580 species' distributions of average radiance values are presented in Appendix 3.

We ended with 477 candidate species from the eBird dataset which matched both the trait taxonomy and the phylogenetic tree taxonomy, and also had complete trait data (Appendix 4).

@@

### Phylogenetic signal
There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1), as it was significantly different than random for 4 out of the 5 indices. Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1). The result for body size-- $k=3.3$ --is among the strongest ever recorded measures of phylogenetic signal in the literature (see REFS) and for $k$ the statistic suggests much stronger signal than expected from Brownian motion. We found weaker signal for behavioral traits compared to morphological traits. For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).

## Model results
### Non-phylogenetic models
The most important predictors of urbanization in the non-phylogenetic global model (Fig. 4) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006). Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032). Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitary (t = -2.90, p = 0.004).

When this global model was dredged and model-averaged, patterns remained similar (Fig. 4). Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors. Additionally, agricultural birds were significantly associated with urbanization (z = 3.40, p = 0.001). Urban birds remained significantly negatively associated with feeding solitary (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).

### Phylogenetically constrained models
The most significant predictors of urbanization in the phylogenetically controlled global model (Fig. 4) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005). Nest generlaism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding species (t = -3.25, p = 0.001) and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.

When the phylogenetic global model was model-averaged, the patterns were similar (Fig. 4). Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index. Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.

### Differences in non-phylogenetic and phylogenetic models
The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the significance of individual predictors (Fig. 4). However, brain residual size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig. 5). Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger parameter estimates) in the phylogenetic model compared with the non-phylogenetic model (Fig. 5). Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig. 4 & Fig. 5). Full model results for each of the four analyses can be found in Appendix 5.

>>>>

### Taxonomic considerations
Given the diverse datasets used in this analysis, there were differing taxonomic authorities followed. The response data from eBird follows the eBird/clements checklist [http://www.birds.cornell.edu/clementschecklist/], while the biological and ecological data follows the working list of Australian birds by BirdLife Australia [http://www.birdlife.org.au/conservation/science/taxonomy], and the phylogenetic tree followed the BirdLife V3 world checklist [http://birdtree.org/taxonomy/]. There were minor differences which existed in common names, scientific names, and recognized species among the three datasets. A taxonomic look-up table was made which combined the three different taxonomies (Appendix 2). Ultimately, a species' response variable, calculated from the eBird database and taxonomic source, was considered for analysis if it matched for both the trait dataset and the phylogenetic tree dataset. If it did not match both, then it was not considered for analysis (e.g., species currently only recognized by eBird/clements such as Western Whistler, Silver-backed Butcherbird, or Paperbark Flycatcher). Additionally, one species (Aleutian Tern) met the criteria for number of observations above, but was not included in the Australian trait data because it is new to Australia and thus not included in the analysis.

# RESULTS
A total of 637,482 checklists resulting in 5,944,819 observations of 580 different species which had > 100 observations were initially investigated for analyses. The least often recorded species was Black-breasted Buttonquail (N=104) while the most recorded species was Australian Magpie (N=158,615). The mean number of observations for a species was 10,250. Plots of all 580 species' distributions of average radiance values are presented in Appendix 3. After the 580 candidate species from the eBird dataset were considered based on complete data following both the trait taxonomy and phylogenetic tree taxonomy (e.g., Appendix 2), we were left with 477 species for analysis. A complete list of the 477 species, in addition to their traits, number of observations, and their urbanization index can be found in Appendix 4.

@@

### Phylogenetic signal
There was a distinct phylogenetic signal in our urbanization index measure (Fig. 3; Table 1), as it was significantly different than random for 4 out of the 5 indices considered. Additionally, there was strong phylogenetic signal for potential predictor traits, especially body size and brain residual (Table 1). The result for body size ($k=3.3$) is among the strongest ever recorded measures of phylogenetic signal in the literature, and for $k$ the statistic suggests much stronger signal than expected from Brownian motion. We generally found weaker signal for behavioral traits compared with morphological traits. For these traits, the estimates of phylogenetic signal were weaker than Brownian motion but significant in most cases (Table 1).

## Model results
### Non-phylogenetic models
The most important predictors of urbanization in the non-phylogenetic global model (Fig. 4; Appendix 5) were feeding habitat generalism (t = 3.40, p = 0.001), breeding habitat generalism (t = 3.27, p = 0.001), clutch size (t = 4.52, p < 0.001), and diet generalism (t = 2.74, p = 0.006). Urban birds were also positively significantly associated with agricultural habitats (t = 2.88, p = 0.004) and migratory status (t = 2.15, p = 0.032). Urban birds were significantly negatively associated with solitary nesting (t = -2.06, p = 0.040), feeding in pairs and flocks (t = -2.24, p =0.025), being granivorous (t = -3.40, p = 0.001), using forested habitats (t = -2.94, p = 0.003), using grassland/shrubland habitats (t = -4.26, p < 0.001), being insectivorous (t = -3.96, p < 0.001), and feeding solitarily (t = -2.90, p = 0.004).

When this global model was dredged and model-averaged, patterns remained similar (Fig. 4; Appendix 5). Clutch size (z = 4.83, p < 0.001), diet generalism (z = 3.84, p < 0.001), breeding habitat generalism (z = 3.09, p = 0.002), and feeding habitat generalism (z = 3.27, p = 0.001) all remained significant predictors. Additionally, agricultural birds were significantly associated with urbanization (z = 3.40, p = 0.001). Urban birds remained significantly negatively associated with feeding solitarily (z = 3.39, p = 0.001), being insectivorous (z = 5.42, p < 0.001), using grassland/shrubland habitats (z = 4.05, p < 0.001), feeding in pairs and flocks (z = 3.16, p = 0.002), being granivorous (z = 3.91, p < 0.001), and using forested habitats (z = 3.30, p = 0.001).

### Phylogenetically constrained models
The most significant predictors of urbanization in the phylogenetically controlled global model (Fig. 4; Appendix 5) were brain residual size (t = 3.58, p < 0.001), feeding habitat generalism ( t = 4.63, p < 0.001), clutch size (t = 3.04, p = 0.002), diet generalism (t = 2.07, p = 0.039), and breeding habitat generalism (t = 2.80, p = 0.005). Nest generlaism (t = -2.06, p = 0.040), range size, (t = -4.38, p < 0.001), cooperative breeding species (t = -3.25, p = 0.001) and species without any known nest aggregation (t = -3.28, p = 0.001) were significantly negatively associated with the urbanization index.

When the phylogenetic global model was model-averaged, the patterns were similar (Fig. 4; Appendix 5). Feeding habitat generalism (z = 7.54, p < 0.001), clutch size (z = 7.31, p < 0.001), body size (z = 3.93, p < 0.001), and brain residual size (z = 2.10, p = 0.036), were significant predictors of urbanization index. Further, species which were cooperative breeders (z = 2.53, p - 0.011), used grassland/shrubland habitats (z = 7.64, p < 0.001), and no known nest aggregation pattern (z = 4.20, p < 0.001) were significantly negatively associated with urbanization.

### Differences in non-phylogenetic and phylogenetic models
The phylogenetic and non-phylogenetic models largely agreed with the relative rank of standardized predictor estimates, and the statistical significance of individual predictors (Fig. 4). However, residual brain size was not significant in either of the non-phylogenetic models, but was significantly associated with urbanization in our phylogenetic models (Fig. 5). Further, diet generalism, feeding habitat generalism, and breeding habitat generalism were relatively more important (larger standardized parameter estimates) in the phylogenetic models compared with the non-phylogenetic models (Fig. 5). Lastly, although not significant, body size was minimally important for the non-phylogenetic model, but substantially more important for the phylogenetically controlled models (Fig. 4 & Fig. 5). Full model results for each of the four analyses can be found in Appendix 5.
